---
title: "ASSIGNMENT 4"
author: "Xin TANG"
date: '2023-07-12'
output:
  word_document: default
  html_document: default
  pdf_document: default
bibliography: bibliography.bib
editor_options: 
  markdown: 
    wrap: 72
---

# Markdown Basics

## Favorite Foods

1.  Dumplings
2.  Ramen noddle
3.  Fried Rice \## Images

![All Cases (Log
Plot)!](/Users/Daisy/Documents/DSC520/completed/assignment04/plots/10-all-cases-log.png)

## Add a Quote

IN God we trust, rest need to bring data

## Add an Equation

Pythagorean Equation $$
a^2 + b^2 = c^2
$$ \## Add a Footnote

[^1]

[^1]: This is a dummy footnote for my first Rmarkdown.

## Add Citations

[\@ R for Everyone] [\@ Discovering Statistics Using R]

'R for Everyone 'Discovering Statistics Using R

# Inline Code

```{r, include=FALSE}
library(ggplot2)
theme_set(theme_minimal())
setwd("~/dsc520")

## Load the `data/r4ds/heights.csv` to

covid_df <- read.csv("data/nytimes/covid-19-data/us-states.csv")
california_df <- covid_df[ which( covid_df$state == "California"), ]
ny_df <- covid_df[ which( covid_df$state == "New York"), ]
florida_df <- covid_df[ which( covid_df$state == "Florida"), ]

```

## NY Times COVID-19 Data

```{r, echo=FALSE}
ggplot(data=florida_df, aes(x=date, group=1)) +
  geom_line(aes(y = cases, colour = "Florida")) +
  geom_line(data=ny_df, aes(y = cases,colour="New York")) +
  geom_line(data=california_df, aes(y = cases, colour="California")) +
  scale_colour_manual("",
                      breaks = c('Florida', 'New York', 'California'),
                      values = c('red', 'blue', 'green')) +
  xlab(" ") + ylab("Cases") + scale_y_log10()

```

## R4DS Height vs Earnings

```{r, echo=FALSE }
setwd('~/dsc520')
heights_df <- read.csv("data/r4ds/heights.csv")
ggplot(heights_df, aes(x=height, y=earn, col= `sex`)) + geom_point()
```

# Tables

## Create a dataframe called characters_df using the following information from LOTR

```{r, echo=FALSE}

name <- c("Aragon", "Bilbo", "Frodo", "Galadriel", "Sam", "Gandalf", "Legolas", "Sauron", "Gollum")
race <- c("Men", "Hobbit", "Hobbit", "Elf", "Hobbit", "Maia", "Elf", "Maia", "Hobbit")
in_fellowship <- c(TRUE, FALSE, TRUE, FALSE, TRUE, TRUE, TRUE, FALSE, FALSE)
ring_bearer <- c(FALSE, TRUE, TRUE, FALSE, TRUE, TRUE, FALSE, TRUE, TRUE)
age <- c(88, 129, 51, 7000, 36, 2019, 2931, 7052, 589)

characters_df <- data.frame(name, race, in_fellowship, ring_bearer, age)
## Knitr Table with Kable
knitr::kable(
  characters_df,
  caption = 'One Ring to Rule Them All'
)
```

## Pandoc Table

| Name   | Race   | In Fellowship? | Is Ring Bearer? |  Age |
|--------|--------|----------------|-----------------|-----:|
| Aragon | Men    | Yes            | No              |   88 |
| Bilbo  | Hobbit | No             | Yes             |  129 |
| Frodo  | Hobbit | Yes            | Yes             |   51 |
| Sam    | Hobbit | Yes            | Yes             |   36 |
| Sauron | Maia   | No             | Yes             | 7052 |

# References

@lander2014r @field2012discovering
