---
title: "Week 9 Logical Regression"
author: "Xin TANG"
date: '2023-08-1'
output:
  pdf_document: default
  
editor_options: 
  markdown: 
    wrap: 72
---

# Logical Regression: surgery and binary data

## Surgery fit model

working with the thoracic surgery data set

1.  Fit a binary logistic regression model to the data set that predicts
    whether or not the patient survived for one year (the Risk1Y
    variable) after the surgery. include a summary()

# Inline Code

```{r, echo=FALSE}
setwd('~/Dsc520')
library(readr)
library(dplyr)
library(ggplot2)
library(coefplot)
library(scales)

# the surgery data
surgery9 <- read.csv('./data/SurgeryData.csv')

#first tried model with all parameters, then figured out the most relavalant predictor, using those to re-build the model. 

Risk3 <- glm(Risk1Yr ~ PRE14 + PRE7 + PRE9 + PRE25, data = surgery9, family = binomial(link = 'logit'))
coefplot(Risk3)
summary(Risk3)

```

2.  From model and chart, the **PRE14** has greatest effect on the
    survival rate

3.  Compute the accuracy of your model, use the dataset to predict the
    outcome variable.

```{r, echo=FALSE}
#prediction result
predict_risk <- predict(Risk3, newdata = surgery9, type='response')

matrix = table(actual = surgery9$Risk1Yr, predict = predict_risk>0.5)
matrix
accuracy1 = (matrix[1,1]/sum(matrix))
accuracy1

```

The **accuracy** of the model is **85%**

## Logical Regression model on binary data.

1.  Fit a logistic regression model to the binary-classifier-data.csv
    dataset

```{r, echo=FALSE}

binary9 <- read.csv('http://content.bellevue.edu/cst/dsc/520/id/resources/binary-classifier-data.csv')
predict1 <- glm(label ~ x + y, data = binary9, family = binomial(link = 'logit'))
coefplot(predict1)
summary(predict1)

```

### What is the accuracy of the logistic regression classifier?

```{r, echo=FALSE}

#predict result
predict_binary <- predict(predict1, newdata = binary9, type='response')

matrix2 = table(actual =binary9$label, predict = predict_binary>0.5)
matrix2

accuracy_binary_model = (matrix2[1,1]+matrix2[2,2])/sum(matrix2)
accuracy_binary_model

```

The **accuracy** of the model is **58%**

## End of Markdown
